### 2016-02-16 paper.tex
Add Nathan's story. Misc improvements.

\documentclass[twocolumn]{article}
\usepackage[T1]{fontenc}
\usepackage[cm-default]{fontspec}
\usepackage{lmodern}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{sectsty}
\usepackage{ctable}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage[numbers]{natbib}
\usepackage{xr}
\usepackage{authblk}

% TODO: wtf does this break things?
%\externaldocument{supplement}

\setromanfont{Adagio_Serif Light}
\setsansfont[BoldFont={Adagio_Sans}]{Adagio_Sans Light}
\allsectionsfont{\normalfont\sffamily\bfseries}

\title{Isolator: accurate and consistent analysis of isoform-level expression in RNA-Seq experiments}

\author[1]{Daniel C. Jones}
\author[2]{Kavitha T. Kuppusamy}
\author[3]{Nathan J. Palpant}
\author[4]{Xinxia Peng}
\author[4]{Michael G. Katze}
\author[5]{Charles E. Murry}
\author[6]{Hannele Ruohola-Baker}
\author[1,7,8]{Walter L. Ruzzo}

%\date{}

\affil[1]{Department of Computer Science and Engineering, University of Washington}
\affil[2]{XXX}
\affil[3]{XXX}
\affil[4]{Department of Microbiology, University of Washington}
\affil[5]{Department of Pathology, University of Washington}
\affil[6]{Department of Biochemistry, University of Washington}
\affil[7]{Department of Genome Sciences, University of Washington}
\affil[8]{Fred Hutchinson Cancer Research Center}

%\affil[1]{Department of Computer Science and Engineering,
%University of Washington, Seattle, WA 98195-2350, USA}
%\affil[2]{Department of Genome Sciences, University of Washington, Seattle, WA
%98195-5065, USA}
%\affil[3]{Fred Hutchinson Cancer Research Center, Seattle, WA 98109, USA}
%\affil[4]{Department of Microbiology, University of Washington, Seattle, WA
%98195-7242, USA}

\begin{document}
\maketitle

\begin{abstract}
    While RNA-Seq has enabled great progress towards the goal of wide-scale
    isoform-level mRNA quantification, short reads have limitations when
    resolving complex or similar sets of isoforms. As a result, estimates of
    isoform abundance carry far more uncertainty than those made at the gene
    level. When confronted with this uncertainty, commonly used methods produce
    estimates that are unstable---small perturbations in the data often produce
    dramatically different results, confounding downstream analysis. We
    introduce a new method, Isolator, which analyzes all samples in an
    experiment in unison using a simple Bayesian hierarchical model. Combined
    with aggressive bias correction, it produces estimates that are
    simultaneously accurate and stable. In a comprehensive comparison of
    accuracy and stability, we show that this property is unique to Isolator. We
    further demonstrate that the approach of modeling an entire experiment
    enables new analyses, which we demonstrate by examining splicing
    monotonicity across several time points in the development of human
    cardiomyocyte cells.
\end{abstract}

\section{Introduction}\label{introduction}

Since its introduction, RNA-Seq has rapidly become a preferred method of
studying gene expression, having proved to be reliable, reproducible, and
increasingly affordable. In principle, RNA-Seq enables measurements of
expression at a scale and resolution than has previously been impractical or
impossible. Rather than relying on an often vaguely defined notion of a gene,
expression of specific isoforms or exons can be measured. In practice this
promise is difficult to realize, with methods designed to detect changes in
splicing rarely concuring \cite{Liu:2014dn}.

Most RNA-Seq procotocols produce short reads (typically 30 to 150 nucleotides,
often paired-end).  Because a short read is often compatible with many isoforms,
the signal produced from sequencing must be deconvoluted, requiring a certain
level of algorithmic sophistication. This problem was tackled early on and with
great success by Cufflinks \cite{Trapnell:2010kd}, and indeed many of the
methods developed in the interim have followed from the same basic model of
RNA-Seq in which transcripts are represented as particular distributions over
possible reads, and the problem is then to infer mixing coefficients (relative
expression values) inducing a mixture model that best explains the data
\cite{Pachter:2011wm}.

We built on this work, implementing a new method called Isolator. Though
it retains the same underlying model of RNA-Seq, we approach the problem
from the perspective that isoform-level estimates possess a degree of
uncertainty that renders maximum likelihood point-estimates alone a
fundamentally inadequate solution to the problem. As the purpose of gene
expression studies is to compare expression between multiple samples,
our goal, beyond maximizing a narrow definition of accuracy, is to
provide a tool that effectively accounts for this uncertainty in a
coherent and reliable manner.

Rather than estimating expression of individual samples, Isolator uses a
hierarchical model of an entire experiment, for example, including all samples
and time points in a time series. In addition to inferring expression values for
the individual samples, we introduce parameters representing condition-wise and
experiment-wise expression and splicing, as well as per-transcript variance
parameters that are shared across conditions. As expression studies often use
only a small number of replicates, shared variance parameters allows us to make
more accurate estimates of biological variability.

An efficient Markov chain Monte Carlo algorithm is used to generate samples over
the parameters of this model, which are saved in an HDF5-based format
\cite{hdf5}. The output from the sampler can then be processed to produce point
estimates, credible intervals, posterior probabilities, and diagnostics. Though
designed to be run on multiple samples concurrently, Isolator can be run as a
conventional tool on individual samples.

The broad effect of this model is to encode the common assumption that
genes tend to be similarly expressed and similarly spliced between
replicates, and to a lesser extent between conditions. In the absence of
sufficient evidence to the contrary, this informative prior shrinks
estimates towards a baseline of no change, producing more conservative
estimates of effect size than would be obtained by using, for example, a
flat Dirichlet prior. This feature of the model is particularly
important for robust treatment of low-expression transcripts. With few reads
there is little power to determine splicing patterns. Considering the
samples in isolation will often produce very different estimates.
Statistical tests based on these estimates, without considering the full
posterior distribution, make approximations by either
disregarding the relative or compositional nature of the data and
assuming independence, or by relying on asymptotic properties that may
not hold when a gene is not deeply sequenced.

In addition to this hierarchical model, we attempt to aggressively correct for
technical effects. The naive model of RNA-Seq assumes fragments are sampled
uniformly at random from transcripts in proportion to their abundance. Early on,
this was shown to be far from an accurate description of real sequencing data
\cite{Hansen:2010if}. Though severe at nucleotide resolution, gene level
expression estimates are only moderately affected by this bias, as it tends to
average out over long transcripts. When one considers expression at the isoform
level, attempting to correct for this nonuniformity becomes vital, as the
difference between two isoforms is often only a few nucleotides.

We previously developed an efficient and versatile model to account for sequence
specific bias that commonly occurs at the ends of fragments \cite{Jones:2012hg}.
In Isolator, we supplemented this with methods accounting for 3' bias, fragment
GC-content bias, and positional bias caused by end-effects of fragmentation.

Isolator is available under a permissive open source license at:
https://github.com/dcjones/isolator

\section{Results}

The convention for evaluating the accuracy of RNA-Seq analysis methods has
consisted chiefly of comparing point estimate accuracy in simulations,
supplemented with comparisons to qPCR or other technologies. Tuning methods
exclusively to maximize accuracy on simulated data risks solving only an
idealized mathematical model of RNA-Seq while disregarding noise, bias, and
nonuniformity. As no gold standard for whole-transcriptome isoform-level
estimation exists, we adopt a multifaceted approach, focusing on measuring
the stability or consistency of estimates, along with accuracy.

Comparisons of gene or transcript expression are sensitive to how one chooses to
measure similarity. Because gene expression typically varies across many orders
of magnitude, measurements like the Pearson correlation, $L_1$ or $L_2$
distance, or root-mean-square error are often dominated by the most highly
expressed genes. Recognizing this, a popular alternative is Spearman's rank
correlation, which conversely can overemphasize the majority of transcripts,
which have no, or very low expression. These issues are discussed in depth by
Lovell et al \cite{Lovell:2015il}, who propose measurements of proportionality
as a more meaningful metric. We adopt their ``proportionality correlation'',
defined as $\text{pcor}(x, y) = 2\text{cov}(\log(x), \log(y)) /
(\text{var}(\log(x)) + \text{var}(\log(y)))$, where is $\text{cov}(x)$ and
$\text{var}(x)$ give the covariance and variance of $x$, respectively.  Similar
to other measures of correlation, proportionality correlation varies from -1 to
1, with 1 indicating perfectly proportionality, and -1 perfect reciprocality.
Zeros are accounted for by converting all measurements to transcripts per
million (TPM) and adding 0.1.
% TODO: add supplementary info about why 0.1 is a reasonable value.

Many methods have been proposed to aid in the analysis of RNA-Seq
experiments. We limit our discussion specifically to those that seek to
estimate the relative abundance a set of known transcripts. These
include, non-exhaustively, Cufflinks \cite{Trapnell:2010kd},
RSEM \cite{Li:2011cb},
eXpress \cite{Roberts:2012dh},
BitSeq \cite{Glaus:2012ek},
Sailfish \cite{Patro:2014jd},
Salmon (a refinement of Sailfish currently in development), and
Kallisto \cite{Bray:2015uj}.

RSEM was used to produce both maximum likelihood and posterior mean estimates,
which we label ``RSEM/ML'' and ``RSEM/PM'', respectively.  This is not an
exhaustive account of such methods, but does represent a wide variety of popular
approaches. When possible, the same read alignments were used for each method,
which notably excludes the alignment-free methods Sailfish, Salmon, and
Kallisto.
% TODO: see supplementary section XXX

\subsection{Agreement with qPCR and spike-in controls}

Spike-in controls and qPCR are far more limited than RNA-Seq, measuring
only the abundance of specific loci in the case of qPCR, or known
proportions of simple artificial transcripts with spike-in controls.
However, both are considered reliable enough to serve as proxy gold
standards for gene-level expression estimates.

We use data generated by the Sequencing Quality Control Consortium (SEQC)
\cite{Consortium:2014kc}, consisting of four reference samples (labeled A, B, C,
and D). The RNA-seq methods were run using approximately 25 million reads per
sample, obtaining estimates that were then compared to qPCR in Figures
\ref{table:taqman} and \ref{table:primepcr}, and the known ERCC spike-in
mixtures in Figure \ref{table:ercc}. In each of these we find that Isolator
produces the highest correlation of the methods compared, though sometimes by a
small margin. Despite two different technologies, and two different sets of
genes, the qPCR benchmarks agree closely, ranking the methods nearly
identically. Cufflinks, eXpress, Salmon, and RSEM/ML typically perform similarly
to each other, with BitSeq, Kallisto, Sailfish, and RSEM/PM trailing slightly.
Results are similar with ERCC spike-in controls, the notable exceptions being
Kallisto and Sailfish, which show higher accuracy, and eXpress, which shows
lower accuracy compared to their relative rankings in the qPCR benchmarks.

\begin{figure*}
\begin{subtable}{0.5\linewidth}
\caption{Comparison of estimates to TaqMan qPCR}
\begin{tabular}[c]{@{}lrrrr@{}}
\toprule\addlinespace
Method & A & B & C & D
\\\addlinespace
\midrule
Isolator & \textbf{0.903} & \textbf{0.908} & \textbf{0.890} &
\textbf{0.883}
\\\addlinespace
eXpress & 0.901 & 0.902 & 0.884 & 0.874
\\\addlinespace
Cufflinks & 0.900 & 0.901 & 0.865 & 0.872
\\\addlinespace
Salmon & 0.899 & 0.897 & 0.882 & 0.871
\\\addlinespace
RSEM/ML & 0.899 & 0.896 & 0.881 & 0.870
\\\addlinespace
Kallisto & 0.893 & 0.885 & 0.874 & 0.860
\\\addlinespace
BitSeq & 0.875 & 0.865 & 0.863 & 0.848
\\\addlinespace
Sailfish & 0.879 & 0.857 & 0.854 & 0.834
\\\addlinespace
RSEM/PM & 0.878 & 0.866 & 0.864 & 0.848
\\\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\label{table:taqman}
\end{subtable}
\begin{subtable}{0.5\linewidth}
\caption{Comparison of estimates to PrimePCR qPCR}
\begin{tabular}[c]{@{}lrrrr@{}}
\toprule\addlinespace
Method & A & B & C & D
\\\addlinespace
\midrule
Isolator & \textbf{0.878} & \textbf{0.866} & \textbf{0.839} &
\textbf{0.851}
\\\addlinespace
Cufflinks & 0.870 & 0.856 & 0.799 & 0.841
\\\addlinespace
eXpress & 0.870 & 0.855 & 0.829 & 0.840
\\\addlinespace
Salmon & 0.866 & 0.852 & 0.826 & 0.836
\\\addlinespace
RSEM/ML & 0.865 & 0.851 & 0.825 & 0.835
\\\addlinespace
BitSeq & 0.840 & 0.821 & 0.802 & 0.813
\\\addlinespace
Kallisto & 0.858 & 0.840 & 0.817 & 0.826
\\\addlinespace
Sailfish & 0.844 & 0.814 & 0.797 & 0.802
\\\addlinespace
RSEM/PM & 0.840 & 0.822 & 0.803 & 0.811
\\\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\label{table:primepcr}
\end{subtable}

\begin{subtable}{0.5\linewidth}
\caption{Comparison of estimates to ERCC spike-in controls}
\begin{tabular}[c]{@{}lrrrr@{}}
\toprule\addlinespace
Method & A & B & C & D
\\\addlinespace
\midrule
Isolator & \textbf{0.980} & \textbf{0.979} & \textbf{0.981} &
\textbf{0.983}
\\\addlinespace
Salmon & 0.976 & 0.975 & 0.978 & 0.979
\\\addlinespace
Kallisto & 0.972 & 0.972 & 0.973 & 0.976
\\\addlinespace
Sailfish & 0.970 & 0.969 & 0.969 & 0.972
\\\addlinespace
Cufflinks & 0.967 & 0.969 & 0.972 & 0.974
\\\addlinespace
RSEM/PM & 0.943 & 0.949 & 0.944 & 0.949
\\\addlinespace
RSEM/ML & 0.941 & 0.948 & 0.945 & 0.951
\\\addlinespace
BitSeq & 0.940 & 0.949 & 0.943 & 0.946
\\\addlinespace
eXpress & 0.931 & 0.939 & 0.935 & 0.942
\\\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\label{table:ercc}
\end{subtable}
\begin{subtable}{0.5\linewidth}
\caption{Consistency of SEQC sample estimates}
\begin{tabular}[c]{@{}lrr@{}}
\toprule\addlinespace
Method & $c$ vs $0.75a + 0.25b$ & $d$ vs $0.25a + 0.75b$
\\\addlinespace
\midrule
Isolator & \textbf{0.978} & \textbf{0.978}
\\\addlinespace
BitSeq & 0.967 & 0.967
\\\addlinespace
RSEM/PM & 0.968 & 0.967
\\\addlinespace
Sailfish & 0.932 & 0.925
\\\addlinespace
RSEM/ML & 0.922 & 0.919
\\\addlinespace
Salmon & 0.916 & 0.914
\\\addlinespace
Kallisto & 0.907 & 0.902
\\\addlinespace
eXpress & 0.903 & 0.899
\\\addlinespace
Cufflinks & 0.870 & 0.916
\\\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\label{table:seqctranscripts}
\end{subtable}

\caption{
Benchmarks measuring the accuracy and consistency of estimates made from 
Sequencing Quality Control (SEQC) dataset using a variety of methods.
All numbers are reported in proportionality correlation.
(a) correlation between gene level quantification
of 806 genes using TaqMan qPCR and RNA-Seq quantification.
(b) correlation between gene level quantification
of 18,353 genes using PrimePCR qPCR and RNA-Seq quantification.
(c) correlation between known proportions of 92
ERCC spike-in controls and RNA-Seq quantification.
(d) correlation between transcript-level estimates
for the mixed SEQC samples C and D and weighted averages of estimates for A
and B, corresponding to the mixture proportions for C and D. Sample C consists of
$3/4$ A and $1/4$ B and sample D is $1/4$ A and $3/4$ B. Consistent estimates of
transcript expression in A, B, C, and D should agree with these mixture
proportions.}
\end{figure*}

\subsection{Estimation consistency}\label{estimation-consistency}

The four SEQC samples consist of two commercially available reference RNA
samples, labeled $A$ and $B$, and two synthetic samples formed by mixing $A$ and
$B$ in specific proportions. Sample $C$ is composed of 75\% $A$ and 25\% $B$,
while sample $D$ is 25\% $A$ and 75\% $B$. This design allows us to directly
measure expression consistency. If we estimate transcript abundance for each
sample, producing estimates $a, b, c, d$, then for these estimates to be in
agreement we should expect $c$ to be approximated by $0.25a + 0.75b$, and
similarly $d$ by $0.75a + 0.25b$. To capture the notion of consistency, we
measured the closeness of this approximation using proportionality correlation
(Figure \ref{table:seqctranscripts}).

Results from RSEM are particularly informative. We generated maximum likelihood
and posterior mean estimates using the same aligned reads. Merely switching the
estimate from the former to the latter increased the correlation by
approximately 5\%. This stands in contrast to comparisons to qPCR which suggest
reduced accuracy of posterior mean estimates, demonstrating that consistency and
accuracy, while related, are two separate axis of comparison.

Estimates made at the gene level (Supplementary Table \ref{table:seqcgenes})
agree with these results, but are of uniformly higher correlation and show a
much smaller gap between posterior mean and maximum likelihood estimate.
Methods reporting posterior mean estimates (Isolator, BitSeq, RSEM/PM) again
show the highest correlation and are in very close agreement.

\subsection{Batch effects}\label{batch-effects}

\begin{figure*}
\includegraphics[width=\textwidth]{batch-effect-figure.pdf}
\caption{
\textbf{a} A heatmap showing pairwise proportionality correlation between
samples sequenced on two flowcells each at five sites, from centrally prepared
libraries. Flowcells are numbered arbitrarily 1 or 2 and sequencing sites are
abbreviated with three letter codes: Australian Genome Research Facility (AGR),
Beijing Genome Institute (BGI), Cornell University (CNL), Mayo Clinic (MAY), and
Novartis (NVS). \textbf{b} The absolute change in correlation induced by
enabling bias correction for methods that support it. For clarity this plot
excludes points for BitSeq estimates of "MAY 2", as bias correction has an
extremely detrimental effect on these. Mean improvement in correlation was 0.008
with Salmon, 0.007 with Cufflinks, 0.006 with Isolator, 0.003 with Kallisto,
0.002 with eXpress, and -0.164 with BitSeq.
}

\label{fig:batcheffects}
\end{figure*}

We further examined the question of consistency by comparing the same
samples sequenced on different flow cells and at different sites, again
using data from the SEQC. We used 10 samples from the larger SEQC
dataset, each consisting of a single lane from two separate
flowcells sequenced at five different sites, all on Illumina HiSeq 2000
instruments, and compared pairwise agreement between estimates from these
samples (Figure \ref{fig:batcheffects}a).

These correlations are somewhat smaller than those in Table
\ref{table:seqctranscripts}, largely because of shallower sequencing: in Table
\ref{table:seqctranscripts} five flowcell lanes were combined, totaling
\textasciitilde{}25 millions reads per sample, here only one lane
(\textasciitilde{}5 million reads) was used. Nevertheless, we again see
posterior mean estimates (as in Isolator, BitSeq, and RSEM/PM) with
significantly higher agreement between pairs.

When accounting for batch effects, bias correction methods can have a
significant impact. To examine the efficacy of bias correction, we disabled this
functionality on those methods that support it, repeated the experiment, and
measured the change in pairwise correlation (Figure \ref{fig:batcheffects}b). We
see that bias correction is largely beneficial for each method. Isolator,
Salmon, and Cufflinks show similar improvements with bias correction, though in
a few cases Cufflinks slightly decreases agreement. Kallisto show a consistent,
but very slight improvement. BitSeq's bias correction sometimes has a very
positive effect, but other times a negative or even disastrous effect. Estimates
from the ``MAY 2'' samples in particular had far worse agreement with other
samples with bias correction enabled. Here bias correction has dramatic and
detrimental effects on some number of transcripts, driving the correlation to
negative numbers.

While this benchmark is informative, it should be considered a lower bound on
the batch effects and bias found in real experiments: these samples were
sequenced at different labs, but using the same instrument model (HiSeq 2000)
and from libraries prepared centrally. The fact that technical effects were so
clearly present even under such controlled circumstances suggests that
reproducibility should not be taken for granted in RNA-Seq experiments,
particularly when it comes to batch effects.

\subsection{Accuracy in simulated data}\label{accuracy-in-simulated-data}

To more directly demonstrate Isolator's capacity to produce more accurate
estimates by modeling entire experiments, we simulated a simple RNA-Seq
experiment consisting of two conditions each with three replicates. Expression
values were generated from a two-component log-normal model with parameters fit
to Cufflinks estimates of data from Kuppusamy et al \cite{Kuppusamy:2015ey}. We
generated 10 million, 100nt paired-end simulated RNA-Seq reads per sample using
rlsim \cite{Sipos:2013tz}.

In this benchmark, Isolator significantly improves on existing methods (Table
\ref{table:rlsim}). It has the unique advantage of sharing information between
replicates. With the assumption that genes tend to be similarly spliced between
replicates, we are able to more effectively resolve transcript expression in
complex loci. Kallisto, Salmon, RSEM/ML, and Cufflinks all perform very
similarly. While comparisons to qPCR showed eXpress and Cufflinks with a
significant advantage over RSEM/ML, Kallisto, and Bitseq, effective bias
correction is less important in simulated data, plausibly explaining the
difference. Although rlsim models some forms of bias, technical effects in
RNA-Seq are not entirely understood, so these are likely understated in its
model.

\begin{table}
\label{table:rlsim}
\begin{tabular}[c]{@{}ll@{}}
\toprule\addlinespace
Method & Correlation
\\\addlinespace
\midrule
\textbf{Isolator} & \textbf{0.919}
\\\addlinespace
Kallisto & 0.887
\\\addlinespace
Salmon & 0.886
\\\addlinespace
RSEM/ML & 0.881
\\\addlinespace
Cufflinks & 0.881
\\\addlinespace
eXpress & 0.825
\\\addlinespace
Sailfish & 0.816
\\\addlinespace
RSEM/PM & 0.806
\\\addlinespace
BitSeq & 0.796
\\\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\caption{
Proportionality correlation between ground truth and estimates
produced by each method using simulated RNA-Seq reads.}
\end{table}

\subsection{Agreement between two sequencing
technologies}\label{agreement-between-two-sequencing-technologies}

RNA-Seq reads become more informative with length and quantity. We
compared data from the same sample sequenced twice, once with 300nt
paired-end reads on a MiSeq sequencer and again with 100nt paired-end
reads using a HiSeq 2000, yielding approximately 4.5 million and 11
million reads, respectively. To elucidate the effect of deeper
sequencing, we sampled without replacement smaller subsets of the 11
million HiSeq 2000 reads and measured correlation between estimates made using
these subsamples with estimates from the MiSeq data (Figure \ref{fig:rhemac}).

Isolator shows the highest correlation, but only when sequencing depth
exceeds a few million reads (below which would be unusually shallow
sequencing for an RNA-Seq experiment) where it overtakes eXpress by a small
margin. The outliers in the this test are BitSeq and RSEM/PM. The former
shows extremely high correlation with small numbers of reads, but
contrary to all other methods, correlation rapidly decreases with deeper
sequencing after 1 million reads. When posterior mean estimates are
generated from a model using a uninformative prior, as is the case with
BitSeq, transcripts with few or no reads may produce estimates that are
highly influenced by transcript length, which plausibly explains this
phenomenon. Yet, RSEM/PM does increase in correlation monotonically with
sequencing depth, though the correlation remains very low. Even using
all 11 million reads, RSEM/PM had a correlation of only 0.567.

\begin{figure}
\includegraphics[width=0.5\textwidth]{rhemac-figure.pdf}
\caption{Proportionality correlation between estimates from 4.5 million 300nt
MiSeq reads and progressively larger numbers of HiSeq 2000 reads. For clarity,
correlations for RSEM/PM are excluded as they were much lower than the other
methods.}
\label{fig:rhemac}
\end{figure}

\subsection{Case study: finding monotonic differential splicing}\label{finding-monotonic-differential-splicing}

Splicing has been known to play an important role in devoloping hearts. Specific
splicing changes during maturation have been shown to be conserved across
chicken and mouse, and influenced by expression of CELF and MBNL
\cite{Kalsotra:2008ke}. Recently, stress-induced switching between two mutually
exclusive isoforms of KHK has been shown to be driver of pathological cardiac
hypertrophy \cite{Mirtschink:2015iq}.

We previously used Isolator to analyze splicing dynamics in human cardiomyocyte
cells during maturation \cite{Kuppusamy:2015ey}. By modeling the entire RNA-Seq
experiment and saving samples generated during sampling, Isolator is uniquely
able to compute posterior probabilities corresponding to arbitrarily complex
questions, within the confines of the model. As this experiment consisted of
a number of samples from a number of conditions and time points, it did not
easily lend itself to analysis using pairwise tests.

We instead searched for differential splicing events by computing the
probability of ``monotonic splicing``, which we defined as a splicing change
that that occurs is consistently observed as occuring in the same direction
between immature (H7 cells at day 20 and at day 30 infected with an empty
vector) and mature samples (H7 1 Year, adult heart, and fetal atrium and
ventricle). To compute the posterior probability of monotonic splicing, we
simply count the number of samples from the model in which this pattern is
observed, and divide by the total number of samples generated (350, in this
case).

This approach identified 272 probable (posterior probability > 0.5) differential
splicing events. We chose seven high probability candidates (ACOT9, IMMT, LTBP4,
OBSL1, OSBPL1A, PALM, and TNNT2) to validate with qPCR (Supplementary Section
\ref{cardioqpcr}), by measuring the expression of specific exons predicted a
change in splicing. Five of these candidates (ACOT9, LTBP4, OBSL1, OSBPL1A, and
TNNT2) showed a clear gradient between H7 day20, fetal, and adult tissue.

We analyzed the same data with the other seven methods. Exon-level fold change
estimates generally agree between RNA-Seq methods, though far less so between
RNA-Seq and qPCR, which could be caused by poor gene annotations or yet
unidentified technical effects. Principal component analysis using transcript
expression estimates also demonstrates some revealing differences (Supplementary
Section \ref{cardiopca}).

\subsection{Case study: novel mechanisms of cell fate specification identified
by gene splice variants}

\begin{figure*}
\begin{center}
\includegraphics[width=0.75\textwidth]{palpant/figure.pdf}
\end{center}
\caption{
\textbf{a}
Hierarchical clustering of the dominant splice variant for 84
differentially spliced genes. Blue arrows indicate genes with known function in
endothelial or heart development. Red arrows indicate genes in the class of
receptors that bind VEGF.
\textbf{b}
Two dimensional principle component analysis in mesodermal progenitor
populations showing discrete separation of populations entirely based on
differences in splice variants. 
\textbf{c}
Analysis of gene ontology showing enriched categories found in 84 differentially
spliced genes. (d-e) Heat map analysis 
\textbf{d}
Heat map showing expression of annotated isoforms of NRP1.
\textbf{e}
Read coverage over isoforms of NRP1.
\textbf{f}
Schematic diagram of protein domains encoded by splice variant variants of NRP1.}
\label{fig:mesodermal}
\end{figure*}

Very little is known about the role of splice variants in mediating cell fate
specification even though it has been shown that 92-95\% of genes are
differentially spliced in a cell or tissue specific manner \cite{Wang:2008ea,
Pan:2008jq}. In order to identify novel mechanisms of cell fate specification in
mesoderm development we began by analyzing RNA-seq data to determine differences
in splice variants between human embryonic stem cell (hESC) derived cardiac
progenitor cells (CPCs) and two endothelial lineages, hESC-derived endocardial
endothelial cells (EECs) and hemogenic endothelial cells (HECs) as we have
described previously \cite{Palpant:2015dv}. 84 genes were identified as
differentially spliced (defined here as a > 0.3 shift in relative transcript
abundance with > 0.5 posterior probability) between these populations (Figure
\ref{fig:mesodermal}a).  Principle component analysis indicated that splice
variants, independent of gene expression levels, delineate endothelial
populations relative to the CPCs (Figure \ref{fig:mesodermal}b). Furthermore,
these data indicate that splice variants alone are sufficient to differentiate
these closely related lineages.  Gene ontology analysis of the 84 differentially
spliced genes showed appropriate enrichment for categories associated with
alternative splicing (Figure \ref{fig:mesodermal}c). Upon closer examination, we
identified neuropilin-1 (NRP1) as an excellent candidate molecule that showed
highly distinct expression profiles of splice variants between the cardiac and
hemogenic endothelial lineages (Figure \ref{fig:mesodermal}d-f).

NRP1 plays a versatile set of roles involving regulation of signaling pathways
that mediate angiogenesis, axon guidance, cell survival, migration, and cell
invasion in a wide range of developmental states and tissue types
\cite{Zachary:2011dr}. It is a single pass transmembrane molecule that has
multiple extracellular functional domains that interact with VEGF and/or
semaphorin molecules \cite{Cackowski:2004ja}. NRP1 has been implicated in
endothelial \cite{Lee:2002dh} and cardiac \cite{Gu:2003vq} development but the
specific splice variants involved in mediating these processes has not been
elucidated.  Results shown here indicate that the cardiac splice variant
contains all functional domains involved in binding VEGF and semaphorin 3a
(Figure \ref{fig:mesodermal}d-f). This is consistent with studies showing the
requirement for NRP1 in mediating the activities of VEGF and semaphorins in
cardiac development \cite{Gu:2003vq}. In contrast, we show that splice variants
expressed in endothelial populations include truncated versions that either lack
the semaphorin binding CUB domains and, in the case of HECs, express a secreted
version of NRP1 (Figure \ref{fig:mesodermal}d-f). This suggests that VEGF and
not semaphorin 3a is the primary mechanism of signaling through NRP1 in
endothelial development \cite{Lee:2002dh}. These results provide the first
evidence to our knowledge that NRP1 splice variants may be key determinants of
fate determination in mesoderm patterning.

\section{Discussion}\label{discussion}

To assess the suitability of Isolator we have compared it to large number of
alternative methods using a wide variety of benchmarks, with a particular focus
on estimation consistency or stability. Data generated by the SEQC is ideally
suited to examine the question consistency, and indeed our data agrees with
their results. The SEQC compared Cufflinks and BitSeq and found BitSeq to
produce more consistent estimates. Here we control for alignment method by using
the same alignments for every method (excluding the alignment-free methods
Salmon, Sailfish, and Kallisto), and compare wider variety methods. In this
broader context it becomes clear that this separation is a consequence of the
choice in estimators. Unmoderated maximum likelihood methods, though often
accurate, frequently demonstrate lower consistency than methods that use MCMC
sampling and report posterior means or medians. This should be unsurprising.
Because of the ambiguity of sequencing data, the likelihood functions for
relative expression are often near plateaus. Maximum likelihood estimates are
thus are precarious compared to posterior mean estimates. Neither estimate is
inherently more correct, and robust analysis should make use of the full
posterior distribution, but we argue that posterior mean or median estimates
are more useful in practice because of this property.

Most methods that report posterior mean estimates show seemingly poor accuracy.
Much of the cause of the seemingly reduced accuracy of the MCMC based models is
driven by low-expression transcripts. A transcript with no reads has a maximum
likelihood expression of zero, but likelihood is positive over a range of
non-zero values, so the posterior mean must necessarily also be positive. With
a flat, or ``non-informative'' prior expression values for these low-expression
transcripts are often quite high and highly determined by transcript length.
Isolator avoids this problem by using a model with informative priors: namely
encoding assumptions that most transcripts have low expression and that
expression and splicing tend to be similar between samples. In this way, it is
able to avoid what seems like a tradeoff and produce estimates that are both
consistent and
accurate.

Fully Bayesian methods like the one presented here possess a degree of
subjectivity that sometimes give researchers pause. Although we have
found the results from Isolator to be insensitive to precise values
given to hyperparameters, they are chosen in advance and with a degree
of arbitrariness. What is often ignored is that the alternative, methods
based on unregularized maximum likelihood estimates, in practice often
necessitate a more insidious form of subjectivity: contriving cutoffs to
filter out unreliable point estimates of low expression genes or
transcripts. While the priors in Bayesian methods are purely explicit,
the means by which data was filtered during an analysis is often a form
of off-the-books subjectivity--critical to the results, yet unmentioned
or only alluded to in manuscripts.

\section{Online Methods}\label{online-methods}

\subsection{Hierarchical Modeling of RNA-Seq
experiments}\label{hierarchical-modeling-of-rna-seq-experiments}

To model RNA-Seq experiments we use a hierarchical model consisting of
three levels: samples or replicates, conditions, and the experiment. Any
number of conditions, and any number of replicates within those
conditions. We model condition-wise and experiment-wise expression for
each transcript using a Gamma distribution, parameterized by mean and
shape. Shape, which capture biological variability, are shared across
conditions.

We further parameterize the model to capture splicing. Despite the
compositional nature, for efficiency purposes we use independent Normal
distributions the model splicing rates at the condition and experiment
level, where a splicing rate is an isoforms proportion relative to total
expression of the its gene.

\subsection{Efficient Sampling}\label{efficient-sampling}

The isoform quantification or deconvolution problem is most easily
thought of as an additive mixture model in which the component
distributions (the isoforms) are known, but the mixture coefficients
(isoform abundances) must be inferred. The standard approach to this
problem is ether expectation maximization (EM) if the maximum likelihood
solution is desired, or Gibbs sampling if, as we are, estimating the
posterior distribution. We have taken a different approach and instead
rely on slice sampling.

The trade-offs between slice sampling and Gibbs sampling for the isoform
quantification problem are subtle. Subsequent samples drawn from a slice
sampler are generally less autocorrelated than those drawn from a Gibbs
sampler. As a result, fewer samples need to be generated to adequately
explore a distribution. However, each of these samples is more
computationally expensive to compute, so one might compensate for
autocorrelation in a Gibbs sampler by drawing more samples in less time.
It is not obvious then that one approach is inherently more efficient
than the other.

Because the likelihood function in this problem is simple, and we sample
from the posterior probability directly, without introducing latent
variables, we are afforded some optimization opportunities not available
in a Gibbs sampler. Specifically, we implement the likelihood function
using SIMD (single instruction, multiple data) instructions, which are
available nearly all CPUs from the last decade. This, in combination
with some numerical approximations, allow us to compute the likelihood
function over an order of magnitude faster on conventional hardware,
than a more straightforward implementation. Run time for Isolator is generally
significantly less than that of other sampling based implementations, though
alignment-free, maximum-likelihood methods like Salmon and Kallisto remain the
most efficient methods (Supplementary Figure \ref{fig:runtime})

\subsection{Correcting for Bias}\label{correcting-for-bias}

Isolator attempts to model and correct for multiple factors that can
conflate the transcript expression estimates.

Perhaps the most prominent source of nonuniformity is sequence specific
bias present at fragment ends. A probable source of this bias is cDNA
synthesis by random priming, which is part of popular RNA-Seq protocols
such as Illumina's TruSeq kit. We have previously published a model that
successfully accounts for effect by training a sparse graphical model
\cite{Jones:2012hg}.

Beyond this random priming bias, there is residual fragment GC-bias,
which may be an artifact of PCR amplification. We observe that fragments
with very high or very low GC-content are less likely to be observed
than expected, even after accounting for priming bias. We model this by
binning fragments by GC content, computing expected and observed
frequencies, and adjusting by their ratio.

Selection of polyadenylated transcripts is a common step mRNA-Seq used
to avoid sequencing introns, partially degraded transcripts, and
ribosomal RNA. This can cause enrichment of reads at the 3' end of
transcripts if only partial transcripts are captured. To model the
effect we fit a one parameter model in which a transcript is truncated
at any position with probability $p$. The probability of observing a
fragment at a position $k$ of length $\ell$, counting from the 3' end,
is then proportional to the probability that the transcript was
\emph{not} truncated before $k + \ell$, which is $(1-p)^{k + \ell}$.
When fit, $p$ is typically quite small, on the order of $10^{-5}$, in
which case this correction has little effect on transcripts shorter that
several kilobases.

Lastly, subtle implications arise from random fragmentation steps in many
protocols. Existing statistical models assume fragments are sampled uniformly
at random from a transcript. However, this does not exactly match the
implications of random fragmentation in which, under an ideal model,
\emph{breakpoints} rather than fragments are introduced uniformly at random. For
a fragment to be observed it must pass size selection and have fallen between
two breakpoints, or one breakpoint and the end of the transcript. Since the ends
of a transcript act as fixed breakpoints, the result is some enrichment of
fragments at either end of the transcript. We compute this enrichment exactly
for various transcript length, and use interpolation to approximate th effect.

\bibliographystyle{plainnat}
\bibliography{references,extra-references}

\end{document}